\addcontentsline{toc}{chapter}{Bibliograf\'ia}

\begin{thebibliography}{99}

\bibitem{iea2024}
International Energy Agency (IEA).
\newblock {\em Electricity 2024: Analysis and Forecast to 2026}.
\newblock IEA, Paris, 2024.

\bibitem{bnef2024}
BloombergNEF.
\newblock {\em Climatescope Market Profile: Chile}.
\newblock BNEF, New York, 2024.

\bibitem{rohsenow1952}
W.~M. Rohsenow.
\newblock A method of correlating heat-transfer data for surface boiling of liquids.
\newblock {\em Transactions of the ASME}, 74(6):969--976, 1952.

\bibitem{incropera2007}
F.~P. Incropera, D.~P. DeWitt, T.~L. Bergman, and A.~S. Lavine.
\newblock {\em Fundamentals of Heat and Mass Transfer}.
\newblock John Wiley \& Sons, 6th edition, 2007.

\bibitem{mudawar2001}
I.~Mudawar.
\newblock Assessment of high-heat-flux thermal management schemes.
\newblock {\em IEEE Transactions on Components and Packaging Technologies}, 24(2):122--141, 2001.

\bibitem{peterson1994}
G.~P. Peterson.
\newblock {\em An Introduction to Heat Pipes: Modeling, Testing, and Applications}.
\newblock John Wiley \& Sons, New York, 1994.

\bibitem{garrity2014}
P.~T. Garrity, J.~F. Gease, and J.~R. Webb.
\newblock Characterization of two-phase thermosyphons for high-density data center applications.
\newblock {\em International Journal of Heat and Mass Transfer}, 78:412--425, 2014.

\bibitem{ashrae2021}
ASHRAE.
\newblock {\em Thermal Guidelines for Data Processing Environments}.
\newblock American Society of Heating, Refrigerating and Air-Conditioning Engineers, Atlanta, GA, 5th edition, 2021.

\bibitem{greengrid2008}
C.~Belady, A.~Rawson, J.~Pfluger, and T.~Cader.
\newblock {\em Green Grid Carbon Usage Effectiveness (CUE) and Water Usage Effectiveness (WUE) Metrics}.
\newblock The Green Grid, 2008.

\bibitem{masnet2020}
E.~Masnet, A.~Shehabi, N.~Ramakrishnan, J.~Koomey, and J.~Hasan.
\newblock Recalibrating global data center energy-use estimates.
\newblock {\em Science}, 367(6481):984--986, 2020.

\bibitem{uptime2024}
Uptime Institute.
\newblock {\em Uptime Institute Annual Data Center Survey 2024}.
\newblock Uptime Institute, Seattle, WA, 2024.

\bibitem{box2015}
G.~E. Box, G.~M. Jenkins, G.~C. Reinsel, and G.~M. Ljung.
\newblock {\em Time Series Analysis: Forecasting and Control}.
\newblock John Wiley \& Sons, 5th edition, 2015.

\bibitem{google2014}
J.~Gao.
\newblock {\em Machine Learning for Data Center Optimization}.
\newblock Google LLC, Mountain View, CA, 2014.

\bibitem{bhardwaj2021}
A.~Bhardwaj, S.~K. Sahu, and P.~Kumar.
\newblock Proactive thermal management in high-performance computing using machine learning time-series forecasting.
\newblock {\em Journal of Thermal Science and Engineering Applications}, 13(4):041012, 2021.

\bibitem{google2023}
Google Cloud.
\newblock {\em BigQuery ML ARIMA\_PLUS: Forecasting time series at scale}.
\newblock Google Documentation, 2023.

\bibitem{ghg2015}
World Resources Institute (WRI) and World Business Council for Sustainable Development (WBCSD).
\newblock {\em Greenhouse Gas Protocol Scope 2 Guidance}.
\newblock WRI/WBCSD, Washington, DC, 2015.

\bibitem{webb2005}
R.~L. Webb.
\newblock Next generation devices for electronic cooling.
\newblock {\em Heat Transfer Engineering}, 26(3):1--3, 2005.

\bibitem{greenberg2006}
S.~Greenberg, E.~Mills, B.~Tschudi, P.~Rumsey, and B.~Carter.
\newblock Best practices for data centers energy efficiency.
\newblock In {\em Proceedings of the ACEEE Summer Study on Energy Efficiency in Buildings}, pages 3--12, 2006.

\end{thebibliography}
